### 2022-06-19 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_metaweb_perspectives@0bc03c93ecf191a0dfc81ea9f4129b7b144260df 🚀

\begin{figure}
\hypertarget{fig:embedding}{%
\centering
\includegraphics{figures/conceptual_2.png}
\caption{Overview of the embedding process. A network (\emph{A}),
represented as its adjacency matrix (\emph{B}), is converted into a
lower-dimensional object (\emph{C}) where nodes, subgraphs, or edges
have specific values (see tbl.~\ref{tbl:methods}). For the purposes of
prediction, this low-dimensional object encodes feature vectors for
\emph{e.g.} the nodes. Embedding also allows to visualize the structure
in the data differently (\emph{D}), much like with a principal component
analysis. From a low-dimensional feature vector , it is possible to
develop predictive approaches. Nodes in an ecological network are
species, for which we can leverage phylogenetic relatedness (\emph{e.g.}
Strydom, Bouskila, et al., 2021) or functional traits to fill the values
of additional species we would like to project in this space (here, I,
J, K, and L) from the embedding of known species (here, A, B, C, and D).
Because embeddings can be projected back to a graph, this allows us to
reconstruct a network with these new species. This approach constitutes
an instance of transfer learning.}\label{fig:embedding}
}
\end{figure}

@@

The popularity of graph embedding techniques in machine learning is more
than the search for structural invariants: graphs are discrete objects,
and machine learning techniques tend to handle continuous data better.
Bringing a sparse graph into a continuous, dense vector space (Xu, 2020)
opens up a broader variety of predictive algorithms, notably of the sort
that are able to predict events as probabilities (Murphy, 2022).
Furthermore, the projection of the graph itself is a representation that
can be learned; Runghen et al. (2021), for example, used a neural
network to learn the embedding of a network in which not all
interactions were known, based on nodes metadata. This example has many
parallels in ecology (see fig.~\ref{fig:prediction}), in which node
metadata can be given by phylogeny or functional traits. Rather than
directly predicting biological rules (see \emph{e.g.} Pichler et al.,
2020 for an overview), which may be confounded by the sparse nature of
graph data, learning embeddings works in the low-dimensional space that
maximizes information about the network structure. This approach is
further justified by the observation, for example, that the
macro-evolutionary history of a network is adequately represented by
some graph embeddings (RDPG; see Dalla Riva \& Stouffer, 2016). In a
recent publication, Strydom, Bouskila, et al. (2021) have used an
embedding (based on RDPG) to project a metaweb of trophic interactions
between European mammals, and transfered this information to mammals of
Canada by using the phylogenetic distance between related clades to
infer the values in the latent sub-space into which the metaweb was
projected. By performing the RDPG step on re-constructed value, this
approach yields a probabilistic trophic metaweb for mammals of Canada
based on knowledge of European species, despite a limited (\(\approx\)
5\%) taxonomic overlap.

Graph embeddings \emph{can} serve as a dimensionality reduction method.
For example, RDPG (Strydom, Bouskila, et al., 2021) and t-SVD (Poisot et
al., 2021) typically embed networks using fewer dimensions than the
original network (the original network has as many dimensions as
species, and as many informative dimensions as trophically unique
species; Strydom, Dalla Riva, et al., 2021). But this is not
necessarilly the case -- indeed, one may perform a PCA (a special case
of SVD) to project the raw data into a subspace that improves the
efficacy of t-SNE (Maaten, 2009). There are many dimensionality
reductions (Anowar et al., 2021) that can be applied to an embedded
network should the need for dimensionality reduction (for example for
data visualisation) arise. In brief, many graph embeddings \emph{can}
serve as dimensionality reduction steps, but not all do, neither do all
dimensionality reduction methods provide adequate graph embedding
capacities.

@@

In this section, we illustrate the embedding of a collection of
bipartite networks collected by \textbf{REF}, using truncated Singular
Value Decomposition (t-SVD). The code to reproduce this example (as a
Jupyter notebook) is available as supplementary material. The resulting
(binary) metaweb \(\mathcal{M}\) has 2131 interactions between 206
parasites and 121 hosts (the same metaweb was used for the illustration
of interaction inference through deep learning in Strydom, Catchen, et
al., 2021), and its adjacency matrix has full rank (\emph{i.e.} it
represents a space with 121 dimensions). All analyses were done using
Julia (Bezanson et al., 2017) version 1.7.2, \emph{Makie.jl} (Danisch \&
Krumbiegel, 2021), and \emph{EcologicalNetworks.jl} (Poisot et al.,
2019).

fig.~\ref{fig:illustration}

\begin{figure}
\hypertarget{fig:illustration}{%
\centering
\includegraphics{figures/illustration.png}
\caption{Need to add text here}\label{fig:illustration}
}
\end{figure}

@@

@@

\leavevmode\hypertarget{ref-Strydom2021FooWeb}{}%
Strydom, T., Bouskila, S., Banville, F., Barros, C., Caron, D., Farrell,
M. J., Fortin, M.-J., Hemming, V., Mercier, B., Pollock, L., Runghen,
R., Riva, G. V. D., \& Poisot, T. (2021). \emph{Food web reconstruction
through phylogenetic transfer of low-rank network representation}.
\url{https://doi.org/10.32942/osf.io/y7sdz}

>>>>

\begin{figure}
\hypertarget{fig:embedding}{%
\centering
\includegraphics{figures/conceptual_2.png}
\caption{Overview of the embedding process. A network (\emph{A}),
represented here as its adjacency matrix, is converted into a
lower-dimensional object (\emph{B}) where nodes, subgraphs, or edges
have specific values (see tbl.~\ref{tbl:methods} for an overview of
methods and their use for species interactions). For the purposes of
prediction, this low-dimensional object encodes feature vectors for
\emph{e.g.} the nodes. Embedding also allows to visualize the structure
in the data differently (see fig.~\ref{fig:illustration}), much like
with a principal component analysis. From a low-dimensional feature
vector, it is possible to develop predictive approaches. Nodes in an
ecological network are usually species (\textbf{C}), for which we can
leverage phylogenetic relatedness (\emph{e.g.}
\textbf{Strydom2022FooWeb?}) or functional traits to fill the values of
additional species we would like to project in this space (here for
nodes I, J, K, and L) from the embedding of known species (here, nodes
A, B, C, and D). Because embeddings can be projected back to a graph,
this allows us to reconstruct a network with these new species
(\textbf{D}). This entire cycle constitutes an instance of transfer
learning, where the transfered information is the representation of
graph \textbf{A} through its embedding.}\label{fig:embedding}
}
\end{figure}

@@

The popularity of graph embedding techniques in machine learning is more
than the search for structural invariants: graphs are discrete objects,
and machine learning techniques tend to handle continuous data better.
Bringing a sparse graph into a continuous, dense vector space (Xu, 2020)
opens up a broader variety of predictive algorithms, notably of the sort
that are able to predict events as probabilities (Murphy, 2022).
Furthermore, the projection of the graph itself is a representation that
can be learned; Runghen et al. (2021), for example, used a neural
network to learn the embedding of a network in which not all
interactions were known, based on nodes metadata. This example has many
parallels in ecology (see fig.~\ref{fig:embedding} \textbf{C}), in which
node metadata can be given by phylogeny or functional traits. Rather
than directly predicting biological rules (see \emph{e.g.} Pichler et
al., 2020 for an overview), which may be confounded by the sparse nature
of graph data, learning embeddings works in the low-dimensional space
that maximizes information about the network structure. This approach is
further justified by the observation, for example, that the
macro-evolutionary history of a network is adequately represented by
some graph embeddings (RDPG; see Dalla Riva \& Stouffer, 2016). In a
recent publication, (\textbf{Strydom2022FooWeb?}) have used an embedding
(based on RDPG) to project a metaweb of trophic interactions between
European mammals, and transfered this information to mammals of Canada
by using the phylogenetic distance between related clades to infer the
values in the latent sub-space into which the metaweb was projected. By
performing the RDPG step on re-constructed value, this approach yields a
probabilistic trophic metaweb for mammals of Canada based on knowledge
of European species, despite a limited (\(\approx\) 5\%) taxonomic
overlap.

Graph embeddings \emph{can} serve as a dimensionality reduction method.
For example, RDPG (\textbf{Strydom2022FooWeb?}) and t-SVD (Poisot et
al., 2021) typically embed networks using fewer dimensions than the
original network (the original network has as many dimensions as
species, and as many informative dimensions as trophically unique
species; Strydom, Dalla Riva, et al., 2021). But this is not
necessarilly the case -- indeed, one may perform a PCA (a special case
of SVD) to project the raw data into a subspace that improves the
efficacy of t-SNE (Maaten, 2009). There are many dimensionality
reductions (Anowar et al., 2021) that can be applied to an embedded
network should the need for dimensionality reduction (for example for
data visualisation) arise. In brief, many graph embeddings \emph{can}
serve as dimensionality reduction steps, but not all do, neither do all
dimensionality reduction methods provide adequate graph embedding
capacities. In the next section (and fig.~\ref{fig:illustration}), we
show how the amount of dimensionality reduction can affect the quality
of the embedding.

@@

In this section, we illustrate the embedding of a collection of
bipartite networks collected by Hadfield et al. (2014), using truncated
Singular Value Decomposition (t-SVD) and RDPG (see
\textbf{Strydom2022FooWeb?} for the full details). Briefly, an RDPG
decomposes a network into two subspaces (left and right), which are
matrices that when multiplied give an approximation of the original
network. The code to reproduce this example is available as
supplementary material (note, for the sake of comparison, that Strydom,
Catchen, et al., 2021 have an example using embedding through PCA
followed by prediction using a deep neural network on the same dataset).
The resulting (binary) metaweb \(\mathcal{M}\) has 2131 interactions
between 206 parasites and 121 hosts, and its adjacency matrix has full
rank (\emph{i.e.} it represents a space with 121 dimensions). All
analyses were done using Julia (Bezanson et al., 2017) version 1.7.2,
\emph{Makie.jl} (Danisch \& Krumbiegel, 2021), and
\emph{EcologicalNetworks.jl} (Poisot et al., 2019).

\begin{figure}
\hypertarget{fig:illustration}{%
\centering
\includegraphics{figures/illustration.png}
\caption{Illustration of an embedding for an host-parasite metaweb,
using Random Dot Product Graphs. \textbf{A}, decrease in approximation
error as the number of dimensions in the subspaces increases.
\textbf{B}, position of hosts and parasites in the first two dimensions
of their respective subspaces. \textbf{C}, predicted interaction weight
from the RDPG based on the status of the species pair in the metaweb.
\textbf{D}, relationship between the position on the first dimension and
parasite generalism.}\label{fig:illustration}
}
\end{figure}

The embedding of the metaweb holds several pieces of information
(fig.~\ref{fig:illustration}). In panel \textbf{A}, we show that the
\(L_2\) loss (\emph{i.e.} the sum of squared errors) between the
empirical and reconstructed metaweb decreases when number of dimensions
(rank) of the subspace increases, with a point of inflection around 25
dimensions. As discussed by Runghen et al. (2021), there is often a
trade-off between the number of dimensions to use (more dimensions are
more computationally demanding) and the quality of the representation.
In this instance, accepting \(L_2 = 500\) as an approximation of the
network means that the error for every position in the metaweb is
\(\approx \left(500/(206\times 121)\right)^{1/2}\). In \textbf{B}, we
show the positions of hosts and parasites on the first two dimensions of
the left and right subspaces. Note that these values largely skew
negative, because the first dimensions capture the coarse structure of
the network: most pairs of species do not interact, and therefore have
negative values. In \textbf{C}, we show the predicted weight
(\emph{i.e.} the result of the multiplication of the RDGP subspaces at a
rank of 25) as a function of whether the interactions are observed,
not-observed, or unknown due to lack of co-occurence. This reveals that
the obserbed interactions have higher predicted weights, although there
is some overlap; the usual approach to identify possible interactions
based on this information would be a thresholding analysis, which is
outside the scope of this manuscript (and is done in the papers cited in
this illustration). Note that the values are not bound to the unit
interval, which emphasizes the need for either scaling or clamping
(although thresholding analyses are insensitive to this choice).
Finally, in \textbf{D}, we show that the embedding, as it captures
structural information about the network, holds ecological information;
indeed, the position of the parasite on the first dimension of the left
sub-space is a linear predictor of its number of hosts.

@@

\leavevmode\hypertarget{ref-Hadfield2014TalTwo}{}%
Hadfield, J. D., Krasnov, B. R., Poulin, R., \& Nakagawa, S. (2014). A
Tale of Two Phylogenies: Comparative Analyses of Ecological
Interactions. \emph{The American Naturalist}, \emph{183}(2), 174--187.
\url{https://doi.org/10.1086/674445}

@@
